---
title: "Acceessing ODATA API"
author: Stephen Sani
output: html_notebook
---
**To access some WHO indicator data stored in GHO via the ODATA API**

Load required packages

```{r}
library(readr)
library(vroom)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(httr)
library(jsonlite)
```

# Accessing indicators stored in GHO via the ODATA API
A function to extract a list of indicators stored in GHO via the API
```{r}
query <- "Indicator"

getSdgIndicators <- function(query){
    base_url = "https://ghoapi.azureedge.net/api/"
    response <- httr::GET(paste0(base_url, query))
    cont <- httr::content(response, as="text")
    data_list <- jsonlite::fromJSON(cont)
    indicators <- do.call(cbind.data.frame, data_list$value)
    indicators
}

indicators <- getSdgIndicators(query)

indicators
```

We find the indicator code for a specific indicator, say the Prevalence of wasted children under 5 years of age
```{r}
all_wasted_indicators <- getSdgIndicators(
    "Indicator?$filter=contains(IndicatorName,%20%27wasted%27)"
) 

wasted_indicators <- getSdgIndicators("Indicator") %>%
    filter(IndicatorName == "Children aged <5 years wasted for age (%)") %>%
    select(IndicatorCode)

wasted_indicators
```

# Extracting some data
A function to extract data for wasted children via the API.
```{r}
indicator_code <- "NUTRITION_564"
getSdgIndicatorData <- function(indicator_code){
    base_url = "https://ghoapi.azureedge.net/api/"
    response = httr::GET(paste0(base_url, indicator_code))
    cont = httr::content(response, as="text")
    data_list = jsonlite::fromJSON(cont)
    indicators = do.call(cbind.data.frame, data_list$value)
    indicators
}
indicator_data = getSdgIndicatorData(indicator_code)

indicator_data
```


We are interested in the data between 2010 and 2020
```{r}
data <- indicator_data %>%
    filter(
        TimeDim %in% c(2010 : 2020)
    ) %>%
    select(Id, TimeDim, ParentLocation, Dim1, NumericValue, DataSourceDim) %>%
    arrange(TimeDim, ParentLocation, Dim1)

data
```

# Summarizing the data
We obtain the number of entries for each ParentLocation disaggregated by TimeDim 
```{r}
parent_location_entries <- data %>%
    group_by(TimeDim, ParentLocation) %>%
    summarise(Entries = n()) %>%
    arrange(TimeDim, ParentLocation)

parent_location_entries
```

# Visualizing the summary data
Let us plot a line graph to show the trend of entries by ParentLocation over the years
```{r}
ggplot(
    parent_location_entries, 
    aes(x = TimeDim, y = Entries, group = ParentLocation, color = ParentLocation)
)+
    geom_line()+
    labs(
        title = "Number of Entries by TimeDim disaggregated by ParentLocation",
        x = "TimeDim (Year)",
        y = "Number of Entries",
        color = "ParentLocation"
    )
```
# Conclusion
Although we are interested in the time period between 2010 to 20120, as of "`r Sys.Date()`", the current date of running this program, data only exist up until 2014 and 2015


